---
title: "P8105_hw2_md3851"
author: "Misbath Daouda"
date: "9/27/2019"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(dplyr)
```

##Problem 1

```{r Mr.Trash Wheel, message=FALSE} 
wheel_data =
read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "Mr. Trash Wheel", skip = 1) %>%
janitor::clean_names() %>%
drop_na(dumpster) %>% 
mutate(
  sports_balls = round(sports_balls),
  sports_balls = as.integer(sports_balls)) %>% 
  select(-x15, -x16, -x17)

wheel_data
```

```{r 2018 precipitation, message=FALSE}
precip2018_data =
read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2018 Precipitation", skip = 1) %>%
janitor::clean_names() %>%
drop_na(total, month) %>%
mutate(
  year = 2018) %>%
select(month, year, total)

precip2018_data
```

```{r 2017 precipitation, message=FALSE}
precip2017_data =
read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2017 Precipitation", skip = 1) %>%
janitor::clean_names() %>% 
drop_na(total, month) %>%
mutate(
  year = 2017) %>%
select(month, year, total)

precip2017_data
```

```{r joined precipitation, message=FALSE}
overall_precip_data = 
  full_join(precip2018_data, precip2017_data, by = c("month", "total", "year")) %>%
  mutate(month = month.name[month])

overall_precip_data

```

The `wheel_data` contains monthly information on overall amounts of trash collected at specific dumpsters in Inner Harbor in Baltimore. It also includes quantities of individual items found in the trash, such as `plastic_bottles`, `grocery_bags`, and `cigarette_buds`. It contains `r nrow(wheel_data)` observations. In 2017, the median number of sports balls found in a dumpster was `r median(pull(filter(wheel_data, year == 2017), sports_balls))`. 

The `overall_precip_data` contains monthly information on quantities (in inches) of precipitation in Inner Harbor in Baltimore. It contains `r nrow(overall_precip_data)` observations. Based on the available data, the total precipitation in 2018 was `r sum(pull(filter(overall_precip_data, year == 2018), total))` inches. 

##Problem 2

```{r pols, message=FALSE}
pols_data = read_csv("./data/pols-month.csv") %>%
  janitor::clean_names() %>%
  separate(mon, into = c("year", "month", "day"), sep="-") %>%
  mutate(
  year = as.integer(year),
  month = as.integer(month),
  day = as.integer(day),
  month = month.name[month]) %>%
  mutate(president = recode(prez_dem, "0" = "gop", "1" = "dem")) %>%
  select(-day, -prez_dem, -prez_gop)

pols_data
  
```

```{r snp, message=FALSE}
snp_data = read_csv("./data/snp.csv") %>%
  janitor::clean_names() %>%
  separate(date, into = c("day", "month", "year"), sep="/") %>%
  mutate(
  year = as.integer(year),
  month = as.integer(month),
  day = as.integer(day),
  month = month.name[month]) %>%
  select(year, month, -day, close)

snp_data
```

```{r unemp, message=FALSE}
unemp_data = read_csv("./data/unemployment.csv") %>%
  janitor::clean_names() %>%
  pivot_longer(
  jan:dec,
  names_to = "month",
  values_to = "unemployment"
  ) %>%
  mutate(
  month = recode(month, "jan" = "January", "feb" = "February", 
                        "mar" = "March", "apr" = "April",
                        "may" = "May", "jun" = "June", 
                        "jul" = "July", "aug" = "August", 
                        "sep" = "September","oct" = "October", 
                        "nov" = "November", "dec" = "December"))

unemp_data
```

```{r combined data, message=FALSE}
partial_538_data = 
  left_join(pols_data, snp_data, by = c("year", "month"))
overall_538_data = 
  left_join(partial_538_data, unemp_data, by = c("year", "month")) 

overall_538_data
```

The `pols_data` contains information on the number of national politicians who are democratic or republican on a monthly basis. It contains `r nrow(pols_data)` observations for the years `r range(pull(pols_data, year))`. In June 2015, there were `r pull(filter(pols_data, year == 2015, month == "June"), gov_gop)` governers from the republican party versus `r pull(filter(pols_data, year == 2015, month == "June"), gov_dem)` from the democratic party. 

The `snp_data` contains `r nrow(snp_data)` observations of the Standard & Poor’s stock market index (S&P), which is often used as a representative measure of stock market as a whole. It provides this information for the years `r range(pull(snp_data, year))`. Over this period, the S&P stock took on values as low as `r round(min(pull(snp_data, close)))` and as high as `r round(max(pull(snp_data, close)))`. 

The `unemp_data` contains monthly and yearly unemployment percentages for the years `r range(pull(unemp_data, year))` and includes `r nrow(pols_data)` observations. The average unemployment rate over this period was `r mean(pull(unemp_data, unemployment), na.rm = TRUE)`%.

These three datasets were pulled together to create the `overall_538_data`, which provides the information described above for the years `r range(pull(overall_538_data, year))` and contains `r nrow(overall_538_data)` observations.


##Problem 3

```{r names, message=FALSE}
names_data = read_csv("./data/Popular_Baby_Names.csv") %>%
janitor::clean_names() %>%
distinct() %>%
mutate(
    gender = str_to_lower(gender),
    ethnicity = str_to_lower(ethnicity),
    childs_first_name = str_to_lower(childs_first_name)) %>%
mutate(
    ethnicity = replace (ethnicity, ethnicity == "black non hisp", "black non hispanic"),
    ethnicity = replace (ethnicity, ethnicity == "white non hisp", "white non hispanic"),
    ethnicity = replace (ethnicity, ethnicity == "asian and paci", "asian and pacific islander"))

names_data
```

The table below shows the rank in popularity of the name “Olivia” as a female baby name over time.
```{r olivia}
olivia_data = filter(names_data, childs_first_name == "olivia") %>%
    select(-gender, -childs_first_name, -count) %>%
    pivot_wider(
    names_from = "year_of_birth", 
    values_from = "rank"
    )

olivia_data
```

The table below shows the most popular names among male children over time.
```{r boys}
boys_data = filter(names_data, gender == "male", rank == "1") %>%
    select(-gender, -count, -rank) %>%
    pivot_wider(
    names_from = "year_of_birth", 
    values_from = "childs_first_name"
    )

boys_data
```

The plot below shows the number of children with a name against the rank in popularity of that name, among white non-hispanic children born in 2016. 
```{r scatter_plot}
scatter_plot = filter(names_data, gender == "male", ethnicity == "white non hispanic", year_of_birth == "2016") %>%
  ggplot(aes(x = rank, y = count)) + geom_point()

scatter_plot 

```

